### 2022-08-16 analysis.qmd
wip

---
title: "Analysis"
format: html
---

```{r}
# Packages
library(scales)
library(patchwork)
library(ggh4x)
library(brms)
library(ggstance)
library(bayestestR)
library(tidyverse)

# Plotting theme
theme_set(
  theme_linedraw(
    base_size = 9,
    # base_family = Font
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.75))
    )
)

# For ordering outcomes
oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences"
)

# Directories for files
dir.create("models", FALSE)
dir.create("out", FALSE)
```

## Prepare data

```{r}
# Load and merge processed data
d_fb <- read_csv("data/fb.csv")
d_outcome <- read_rds("data/outcome.rds") %>% 
  mutate(outcome = factor(outcome, levels = oo))
d_pop <- read_rds("data/population.rds")

# Convert fb metrics into country-year proportions
d_fb <- d_fb %>% 
  left_join(d_pop) %>% 
  mutate(
    dau = dau / pop,
    mau = mau / pop
  )

# Merge
d <- left_join(d_fb, d_outcome) %>% 
  mutate(age = factor(age, levels = c("13-34", "35+")))
```

## Data figure

```{r}
# Figure of all variables over time
p_a <- d %>% 
  select(year:mau) %>% 
  pivot_longer(c(dau, mau), names_to = "outcome", values_to = "val") %>% 
  mutate(outcome = if_else(outcome == "dau", "Daily FB users", "Monthly FB users")) %>% 
  ggplot(aes(year, val, col = age, group = interaction(age, country))) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    "Percentage of population", 
    expand = expansion(.03),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2), 
    minor_breaks = seq(2009, 2019, by = 2)
  ) +
  geom_line(size = .25) +
  facet_grid(age~outcome) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  )

p_b <- d %>% 
  select(-c(mau, dau)) %>% 
  # Exclude rows where FB data existed but not GWP data
  drop_na(val) %>% 
  # Actual lowest age in GWP
  mutate(age = factor(age, labels = c("15-34", "35+"))) %>% 
  ggplot(aes(year, val, col = age, group = interaction(age, country))) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    "Value", 
    expand = expansion(.01),
    breaks = extended_breaks()
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2), 
    minor_breaks = seq(2009, 2019, by = 2)
  ) +
  geom_line(size = .25) +
  facet_grid(age~outcome, labeller = as_labeller(~str_replace(., "_", " "))) +
  theme(legend.position = "none")

p_a / p_b
ggsave("out/Fig-data.png", width = 16, height = 10, units = "cm", dpi = 400)
```

## Models

```{r}
# Set global contrasts
options(contrasts = c(unordered = "contr.sum", ordered = "contr.poly"))

# Prepare data
dm <- d %>% 
  # Centering predictors
  # Grand mean centering so both components are zero-centered
  mutate(across(c(dau, mau), ~. - mean(., na.rm = TRUE))) %>% 
  group_by(country) %>%
  mutate(
    # Between
    dau_cb = mean(dau, na.rm = TRUE),
    mau_cb = mean(mau, na.rm = TRUE),
    # Within
    dau_cw = dau - dau_cb, 
    mau_cw = mau - mau_cb
  ) %>% 
  ungroup() %>% 
  # Center year so that intercepts are at 2014
  mutate(year = year - 2014) %>% 
  # Remove rows where predictor existed but outcome didn't
  drop_na(val)

# Model formula
model_dau <- bf(
  val | se(se, sigma = TRUE) ~
    (year + dau_cw) * age + dau_cb +
    ((year + dau_cw) * age | country),
  family = gaussian()
)
model_mau <- bf(
  val | se(se, sigma = TRUE) ~
    (year + mau_cw) * age + mau_cb +
    ((year + mau_cw) * age | country),
  family = gaussian()
)

fits <- dm %>% 
  select(outcome, year, country, age, contains("_c"), val, se) %>% 
  group_by(outcome) %>% 
  nest() %>% 
  mutate(
    dau = map2(
      data, outcome,
      ~brm(
        model_dau,
        cores = 8, chains = 4, threads = 2,
        backend = "cmdstanr",
        adapt_delta = .95,
        iter = 4000,
        data = .x,
        file = str_glue("models/brm-dau-{.y}")
      )
    ),
    mau = map2(
      data, outcome,
      ~brm(
        model_mau,
        cores = 8, chains = 4, threads = 2,
        backend = "cmdstanr",
        adapt_delta = .95,
        iter = 4000,
        data = .x,
        file = str_glue("models/brm-mau-{.y}")
      )
    )
  )
```

```{r}
# Calculate contrasts for each outcome
h_dau <- c(
  "Average" = "dau_cw = 0",
  "13-34" = "dau_cw + dau_cw:age1 * -1 = 0",
  "35+" = "dau_cw + dau_cw:age1 * 1 = 0",
  "Between-country" = "dau_cb = 0"
)
h_mau <- c(
  "Average" = "mau_cw = 0",
  "13-34" = "mau_cw + mau_cw:age1 * -1 = 0",
  "35+" = "mau_cw + mau_cw:age1 * 1 = 0",
  "Between-country" = "mau_cb = 0"
)
coefs <- fits %>% 
  ungroup() %>% 
  mutate(h = if_else(predictor == "dau", list(h_dau), list(h_mau))) %>% 
  mutate(
    world = map2(
      fit, h,
      ~hypothesis(.x, .y) %>% 
        .[["hypothesis"]] %>% 
        tibble()
    ),
    country = map2(
      fit, h,
      ~hypothesis(.x, .y, scope = "coef", group = "country") %>% 
        .[["hypothesis"]] %>% 
        tibble()
    )
  ) %>% 
  select(-fit, -h)

coefs <- coefs %>% 
  pivot_longer(c(world, country)) %>% 
  unnest(value) %>% 
  mutate(country = if_else(name == "country", as.character(Group), "World")) %>% 
  mutate(country = fct_relevel(country, "World")) %>% 
  # Between country contrast not country-specific!
  filter(!(Hypothesis == "Between-country" & country != "World"))
coefs %>%   
  filter(Hypothesis %in% c("13-34", "35+")) %>% 
  mutate(
    panel_order = interaction(predictor, outcome, country),
    panel_order = fct_reorder(panel_order, Estimate)
  ) %>% 
  ggplot(aes(Estimate, panel_order, col = Hypothesis)) +
  scale_color_brewer("Age", palette = "Dark2") +
  geom_vline(xintercept = 0, lty = 2, size = .25, col = "grey50") +
  geom_pointrangeh(
    aes(xmin = CI.Lower, xmax = CI.Upper),
    size = .25, fatten = 2
    ) +
  facet_grid2(predictor~outcome, independent = "all", scales = "free") +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    legend.position = "none"
    )
```

```{r}
coefs %>% 
  filter(Hypothesis == "Between-country")
```